---
title: "Course 5 Week 2 Project"
author: "Sarah M. Duyos"
date: "January 21, 2017"
output: html_document
---

## 1. Reading in the dataset and/or processing the data

```{r}
data <- read.csv("activity.csv")
```
## 2.Histogram of the total number of steps taken each day

```{r}
totalsteps <- tapply(data$steps, data$date, sum, na.rm = TRUE)
hist(totalsteps, xlab="steps/day", ylab = "number observations", main = "Frequency of Steps per Day", breaks = 10)
```


## 3.Mean and median number of steps taken each day

### Mean
```{r}
mean(totalsteps)
```

### Median

```{r}
median(totalsteps)
```

## 4.Time series plot of the average number of steps taken

```{r}
stepsintervalmean <- tapply(data$steps, data$interval, mean, na.rm = TRUE)
plot(stepsintervalmean, type = "l", xlab = "time", ylab = "average steps", main = "Average of steps by 5-minute intervals")
```

## 5. The 5-minute interval that, on average, contains the maximum number of steps

```{r}
pos <- match(max(stepsintervalmean), stepsintervalmean)
names(stepsintervalmean)[pos]
```

## 6.Code to describe and show a strategy for imputing missing data

### Use average of 5-minutes interval over all days

```{r}
data1 <- data
for (rowcnt in 1:nrow(data1)) 
{
  row = data1[rowcnt, ]
  if (is.na(row$steps)) 
  {
    data1[rowcnt, "steps"]= stepsintervalmean[match(row$interval,names(stepsintervalmean))]
  }
}
```

## 7.Histogram of the total number of steps taken each day after missing values are imputed

```{r}
stepstotal1 <- tapply(data1$steps, data1$date, sum)
hist(stepstotal1, xlab = "time", main = "Total steps per day after imputing values")
```


## 8.Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}
alldays <- weekdays(as.Date(data1$date))
weekends <- (alldays == "Saturday" | alldays == "Sunday")
alldays[weekends] <- "weekend"
alldays[!weekends] <- "weekday"
alldays <- as.factor(alldays)
data1$days <- alldays
dataweekend <- subset(data1, days == "weekend")
dataweekday <- subset(data1, days == "weekday")
stepinterval1we <- tapply(dataweekend$steps, dataweekend$interval, mean)
stepinterval1wd <- tapply(dataweekday$steps, dataweekday$interval, mean)

par(mfrow = c(2, 1))
plot(stepinterval1we, type = "l", xlab = "time", ylab = "number steps", main = "Average number steps (5-min interval) Weekend")
plot(stepinterval1wd, type = "l", xlab = "time", ylab = "number steps", main = "Average number steps (5-min interval) Weekday")
```

